---
title: "MATH 513 Practical Presentation"
author: "10570155, 10696253, 10701983"
date: "10/12/2020"
output: 
  beamer_presentation: 
    theme: Boadilla
    colortheme: beaver
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- Samsung and Apple as competing businesses
- Flagship phones chosen
  + S20FE
  + iPhone12
  + provides freshest data
- S20 was chosen for demonstration purposes

__Tools Utilised__

- RTweet
- Rstudio
- Twitter Developer API

```{r, echo=FALSE, out.width="15%", fig.align = "center", fig.show='hold', fig.cap=""}

knitr::include_graphics(
  c("images/logos/apple_logo.png",
    "images/logos/samsung_logo.jpg",
    "images/logos/rstudio_logo.png", 
    "images/logos/twitter_logo.png")
)
```


## Research

Examined the key hashtags that were being used for the phones chosen

- Checked Samsung's Twitter feed
- Checked Tim Cook's Twitter feed
  + Apple's Twitter contains no tweets
- Checked for number and age of tweets using a simple Twitter search for several hashtags
- Avoided selecting individuals or companies so we can get a wide range of opinions for a more rounded data set


## Issues and overcoming them

- Date refinement on Twitter Premium API functions only
  + Dev suite allowed limited access to full archive and 30 day extraction features
- Possible duplication if we extracted at the same time
  + Full archive and 30 day extraction allowed for refinement of dates and pull tweets back to 2006
- Extraction limited to 18k per 15mins
  + Controlled pull of tweets with date refinement over time as retry on rate limit did not function a expected
- Foreign languages
  + Filter by language of 'en' applied to only pull English tweets
- Different regions with different opinions
  + Region filter didn't work in test so we concluded we would examine this in the analysis stage
- Tweets are pulled from date to backwards to date to
  + Pull tweets in a controlled manner using weeks and days


## Results - Overview of Twitter Data

Time series plots of data pulled from Twitter

- iPhone12
- S20FE
- S20

...

## Results - Sentiment Analysis

__Overview of sentiment analysis__


__Choice of feature to analyse__


## Results - Sentiment Analysis

GRAPH HERE


## Results - Sentiment Analysis

GRAPH HERE



## Improvements & Further Study

__Improvements__

Google Maps API to have region filter

  
...


## Conclusions

- Twitter data provides up to date information for companies to analyse for customer feedback

- Data can provide useful information when analysed correctly



...

## References

- Twitter API Documentation
  + [https://developer.twitter.com/en/docs/twitter-api](https://developer.twitter.com/en/docs/twitter-api)

- R Markdown Cheat Sheet
  + [https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)

- R Markdown Reference Guide
  - [https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)

- MATH513 Lecture and Tutorial Code, Luciana Dalla Valle, University of Plymouth
  - [https://dle.plymouth.ac.uk/course/view.php?id=49628](https://dle.plymouth.ac.uk/course/view.php?id=49628)


