---
title: "MATH 513 Practical Presentation"
author: "10570155, 10696253, 10701983"
date: "10/12/2020"
output: 
  beamer_presentation: 
    theme: Boadilla
    colortheme: beaver
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- Samsung and Apple as competing businesses
- Flagship phones chosen
  + S20
  + iPhone12
  + provides freshest data
- S20FE was chosen for comparison

__Tools Utilised__

- RTweet
- Rstudio
- Twitter Developer API

```{r, echo=FALSE, out.width="15%", fig.align = "center", fig.show='hold', fig.cap=""}

knitr::include_graphics(
  c("images/logos/apple_logo.png",
    "images/logos/samsung_logo.jpg",
    "images/logos/rstudio_logo.png", 
    "images/logos/twitter_logo.png")
)
```


## Research

Examined the key hashtags that were being used for the phones chosen

- Checked Samsung's Twitter feed
- Checked Tim Cook's Twitter feed
  + Apple's Twitter contains no tweets
- Checked for number and age of tweets using a simple Twitter search for several hashtags
- Avoided people so we can get a wide range of opions/tweets for a more rounded data set


## Issues identified

- Unable to filter by region without Google API (paid service)
- Date refinement on Twitter Premium
- Possible duplication if we extracted at the same time
- Extraction capped at 500k tweets using non premium features and past 69 days
- Extraction limited to 18k per 15mins
- Foreign languages
- Different regions with different opinions
- Tweets are pulled from date to backwards to date to


## Overcoming Issues
- Our dev suite allowed for:
  + 15k of full archive extraction
  + 25k of 30 day extraction
- Full archive extraction allows for refinement of dates and pull tweets back to 2006
- Use of premium features avoids some duplication
- RTweet will continue to try until it gets to the number of tweets requested
- Filter by 'en' applied to only pull English tweets
- Region filter didn't work in test so we concluded we would examine this in the analysis stage
- Pull tweets in a 2 month cycle


# Data Extraction

Clean-up required the following:


## Stage 3 - Data Visualisation


...

## Stage 3 - Data Visualisation

...

## Improvements & Further Study

__Improvements__
Access to the Google Maps API would 
  - provide better region filter before pulling the tweets
  - 
  
...

## Next Steps

...

## Conclusions

...