---
title: "The speechanalysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette_speechanalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

|       **Authors**: 10570155, 10696253, 10701983
|       **Date**: 14th January 2021

```{r setup, include=FALSE}
libs <- c('speechanalysis', 'knitr')

invisible(lapply(libs, library, character.only = TRUE))
```

### Usage:
```{r, results=FALSE}
library(speechanalysis)
```

## Donald Trump Speech Data

The `speechanalysis` package comes with an example dataset which can be used to explore the functionality provided. It contains 
data which details 10 speeches made by President Donald Trump throughout September 2020.

This data can be accessed like so:

```{r}
trump_speeches
```

## The Change in Word Frequency Over Time

...

## Frequency of Words per Speech

We can use the `plot_tf_idf(df, n)` function, _where `df` is the data frame to apply to and `n` is the number of 'top' words to 
show,_ to plot a graphical summary of the most frequent words which occur in each speech. This function makes use of the **tf-idf 
index** which computes the frequency of a word, adjusted for how rarely it is used.

```{r, warning=FALSE, message=FALSE, fig.width=12, fig.height=8, out.width="98%"}
plot_tf_idf(trump_speeches, 10)
```

## Zipf's Law

...
