---
title: "The speechanalysis Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette_speechanalysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

|       **Authors**: 10570155, 10696253, 10701983
|       **Date**: 14th January 2021

```{r setup, include=FALSE}
libs <- c('speechanalysis', 'knitr')

invisible(lapply(libs, library, character.only = TRUE))
```

### Usage:
```{r, results=FALSE}
library(speechanalysis)
```

## Donald Trump Speech Data

The `speechanalysis` package comes with an example dataset. It consists of data detailing 10 speeches made by President Donald 
Trump throughout September 2020.

The dataset can be accessed like so:

```{r}
trump_speeches
```

## The Change in Word Frequency Over Time

We can use the `word_frequency(df, c)` function, _where `df` is the data frame to apply to and `c` is a collection of words to check and plot
,_ to plot a graphical summary of the frequency of words over time from the speeches.

```{r, warning=FALSE, message=FALSE, fig.width=10, fig.height=18, out.width="98%"}
word_frequency(trump_speeches, c("america", "biden", "China"))
```

## Frequency of Words per Speech

We can use the `plot_tf_idf(df, n)` function, _where `df` is the data frame to apply to and `n` is the number of 'top' words to 
show,_ to plot a graphical summary of the most frequent words which occur in each speech. This function makes use of the **tf-idf 
index** which computes the frequency of a word, adjusted for how rarely it is used.

```{r, warning=FALSE, message=FALSE, fig.width=10, fig.height=18, out.width="98%"}
plot_tf_idf(trump_speeches, 10)
```

## Zipf's Law

...
